\nwfilename{char.nw}\nwbegindocs{0}% -*- noweb-code-mode: c-mode -*-% ===> this file was generated automatically by noweave --- better not edit it
% $Id: char.nw,v 2.3 2003/06/08 20:38:01 wlux Exp $
%
% Copyright (c) 2003, Wolfgang Lux
% See ../LICENSE for the full license.
%
\subsection{Characters}
The runtime system uses a special kind of nodes to implement
characters because of the different requirements with respect to
showing characters. The function {\Tt{}lit{\_}char\nwendquote} converts a character
into its printable representation (without quotes). The {\Tt{}q\nwendquote} argument
is used to specify the quote character. It will be escaped with a
backslash character in the return value. You may pass {\Tt{}'{\nwbackslash}0'\nwendquote} in
order to avoid escaping quote characters.

\nwenddocs{}\nwbegincode{1}\sublabel{NWerkwu-1dcciR-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWerkwu-1dcciR-1}}}\moddef{char.h~{\nwtagstyle{}\subpageref{NWerkwu-1dcciR-1}}}\endmoddef\nwstartdeflinemarkup\nwenddeflinemarkup
extern const char *lit_char(int c, char q);

\nwnotused{char.h}\nwendcode{}\nwbegindocs{2}\nwdocspar
The standard escape sequences for control characters are defined in
a global table.

\nwenddocs{}\nwbegincode{3}\sublabel{NWerkwu-3fnK27-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWerkwu-3fnK27-1}}}\moddef{char.c~{\nwtagstyle{}\subpageref{NWerkwu-3fnK27-1}}}\endmoddef\nwstartdeflinemarkup\nwprevnextdefs{\relax}{NWerkwu-3fnK27-2}\nwenddeflinemarkup
#include "config.h"
#include <stdio.h>
#include <stdlib.h>
#include "debug.h"
#include "run.h"
#include "heap.h"
#include "stack.h"
#include "eval.h"
#include "threads.h"
#include "cam.h"
#include "trace.h"
#include "char.h"

static const char *asciiTab[] = \{
    "\\\\NUL", "\\\\SOH", "\\\\STX", "\\\\ETX", "\\\\EOT", "\\\\ENQ", "\\\\ACK", "\\\\a",
    "\\\\b",   "\\\\t",   "\\\\n",   "\\\\v",   "\\\\f",   "\\\\r",  "\\\\SO",  "\\\\SI",
    "\\\\DLE", "\\\\DC1", "\\\\DC2", "\\\\DC3", "\\\\DC4", "\\\\NAK", "\\\\SYN", "\\\\ETB",
    "\\\\CAN", "\\\\EM",  "\\\\SUB", "\\\\ESC", "\\\\FS",  "\\\\GS",  "\\\\RS",  "\\\\US"
\};

const char *
lit_char(int c, char q)
\{
    static char buf[8], *cp;
    
    if ( c >= '\\0' && c < ' ' )
        return asciiTab[c];
    else if ( c == 0177 )
        return "\\\\DEL";

    if ( c >= ' ' && c < 0177 )
    \{
        cp = buf;
        if ( c == q || c == '\\\\' )
            *cp++ = '\\\\';
        *cp++ = c;
        *cp   = '\\0';
    \}
    else
        sprintf(buf, "\\\\%d", c);

    return (const char *)buf;
\}

\nwalsodefined{\\{NWerkwu-3fnK27-2}}\nwnotused{char.c}\nwendcode{}\nwbegindocs{4}\nwdocspar
The prelude functions {\Tt{}ord\nwendquote} and {\Tt{}chr\nwendquote} convert between integer
numbers and characters. At present, the character range is limited to
8 bit characters. If {\Tt{}chr\nwendquote} is passed an argument outside of this
range it raises an exception. Future versions will support the full
Unicode character set.

\nwenddocs{}\nwbegincode{5}\sublabel{NWerkwu-3fnK27-2}\nwmargintag{{\nwtagstyle{}\subpageref{NWerkwu-3fnK27-2}}}\moddef{char.c~{\nwtagstyle{}\subpageref{NWerkwu-3fnK27-1}}}\plusendmoddef\nwstartdeflinemarkup\nwprevnextdefs{NWerkwu-3fnK27-1}{\relax}\nwenddeflinemarkup
DECLARE_ENTRYPOINT(__ord);

FUNCTION(__ord)
\{
    int c;
    Node *r;

    EXPORT_LABEL(__ord)
 ENTRY_LABEL(__ord)
    EVAL_RIGID_CHAR(__ord);
    c   = (*sp++)->ch.ch;

#if ONLY_BOXED_OBJECTS
    CHECK_HEAP(int_node_size);
    r       = (Node *)hp;
    r->info = &int_info;
    r->i.i  = c;
    hp     += int_node_size;
#else
    r = mk_int(c);
#endif

    RETURN(r);
\}


DECLARE_ENTRYPOINT(__chr);

FUNCTION(__chr)
\{
    int c;

    EXPORT_LABEL(__chr)
 ENTRY_LABEL(__chr)
    EVAL_RIGID_INT(__chr);
    c   = int_val(sp[0]);
    sp += 1;

    if ( c < 0 || c > 255 )
    \{
        fprintf(stderr, "chr: code %d is out-of-range\\n", c);
        exit(1);
    \}
   RETURN((Node *)(char_table + c));
\}
\nwendcode{}

\nwixlogsorted{c}{{char.c}{NWerkwu-3fnK27-1}{\nwixd{NWerkwu-3fnK27-1}\nwixd{NWerkwu-3fnK27-2}}}%
\nwixlogsorted{c}{{char.h}{NWerkwu-1dcciR-1}{\nwixd{NWerkwu-1dcciR-1}}}%

